---
title: "Homework 13"
author: "Tauqeer Kasam Rumaney"
date: "February 4, 2023"
output:
  pdf_document: default
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Survival analysis

Choose one other provided dataset in the survival package (or any other suitable dataset) which contains time-to-event data. Use the survfit function to compute the Kaplan-Meier estimate of the survival function and plot it. Then choose a variable with 2 or more categories in the dataset and perform a log-rank test. Interpret the results: is there are difference in survival between the groups?

# Load the dataaset

```{r}
library(survival)
```

# Survival analysis with the colon dataaset

```{r}
data <- diabetic
head(data)
?diabetic
```

```{r}
# compute the Kaplan Meier estimate of the survival function
KM <- survfit(Surv(time, status) ~ 1, data = data)

# Look at the results (in table form)
summary(KM)

# Look at the plot:
plot(KM, main = expression(paste("Kaplan-Meier-estimate ", hat(S)(t), " with 95% CI")), ylab = "Survival function", xlab = "Time")

# Stratify this by eye (1=left, 2=right):
str(data$eye)
KM2 <- survfit(Surv(time, status) ~ eye, data = data)
plot(KM2, main = expression(paste("Kaplan-Meier-estimate ", hat(S)(t), " stratified by eye")), ylab = "Survival function", xlab = "Time", lwd=2, col=1:2)
legend(x="topright", col = 1:2, lwd = 2, legend=c("left", "right"))

# perform log-rank test:
logrank <- survdiff(Surv(time, status) ~ eye, data = data)
# results:
logrank

```

#Conclusion: The Chi-Squared test statistic is 3 with 1 degree of freedom and the corresponding p-value is 0.08. Since this p-value is more than 0.05, we don't reject the null hypothesis i.e. we don't have sufficient evidence to say that there is a statistically significant difference in survival between the two groups (left eye and right eye).

#We can see that the survival curves are slightly different, but the log rank test tells us that the difference is not statistically significant.