---
title: "Homework 7"
author: "Tauqeer Kasam Rumaney"
date: "December 5, 2022"
output:
  pdf_document: default
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Preparations

Load KiGGS data:

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
kiggs <- KiGGS03_06
```

## Exercise 1: Multiple imputation

a) In the KiGGS dataset, choose 2 metric variables of your choice.
#PPoint and KreaUrin
b) Compute the correlation between these two variables
```{r}
res0 <- cor.test(kiggs$PPoint, kiggs$KreaUrin, method = "pearson")
res0$estimate
res0$conf.int
res0$p.value
```

c) Perform a multiple imputation using the mice function with the 'pmm' method and otherwise default settings.
```{r}
library(mice)

dat <- data.frame(PPoint = as.numeric(as.character(kiggs$PPoint)), 
                  KreaUrin = as.numeric(as.character(kiggs$KreaUrin)))
table(!is.na(dat$PPoint))
table(!is.na(dat$KreaUrin))

# Pattern of missing values
md.pattern(dat)

# Nice visualization of missing values in the VIM package, can be useful to identify patterns of missing values
library(VIM)
aggr(dat, col = c('navyblue','red'), numbers = TRUE, sortVars = TRUE, labels = names(dat), cex.axis = 0.7, gap = 3, ylab = c("Histogram of missing data", "Pattern"))

# Step 1: Do imputation of missing values, generate 5 imputed datasets
tempData <- mice(dat, m = 5, maxit = 5, meth = 'pmm', seed = 500)
summary(tempData)

# The 5 imputed datasets can be extracted as follows:
completedData1 <- complete(tempData,1)
completedData2 <- complete(tempData,2)
completedData3 <- complete(tempData,3)
completedData4 <- complete(tempData,4)
completedData5 <- complete(tempData,5)

```

d) Compute the correlation again on the multiple imputed datasets.
```{r}
# Step 2: Do the statistical analysis using the "with" function
modelFit1 <- with(tempData, cor(PPoint, KreaUrin, method = "pearson"))
mean(unlist(modelFit1$analyses))
with(tempData, cor.test(PPoint, KreaUrin, method = "pearson"))
modelFit1 <- with(tempData, lm(PPoint ~ KreaUrin))
summary(pool(modelFit1), dfcom = 17638)
```

