---
title: "Homework 5"
author: "Tauqeer Kasam Rumaney"
date: "November 20, 2022"
output:
  html_document: default
  pdf_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

## Exercise 1: Probability distributions

Explore the functions *rnorm*, *rt*, *runif*, *rbinom* in R that allow you to generate random numbers from the normal, t-, uniform, and binomial distribution. Compute them with different values, and inspect histograms to visualize their distribution.
```{r}
x <- rnorm(1000, mean=350, sd=15)
hist(x, probability=TRUE)

s <- rt(10,2)
t <- rt(10,1)
hist(s)
hist(t)

z <- runif(20, min = -1, max = 1)
hist(z, probability=TRUE)

y <- rbinom(10, size=100, prob=0.75)
hist(y, probability=TRUE)

```


## Exercise 2: Odds ratio

In the KiGGS dataset:

```{r}
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06
```


a) Compute the proportion of mothers that had hypertension during pregnancy. Use the variable 'e0155' which has values "Ja" (yes), "Nein" (No) and "Weiß nicht" (don't know).
#202
```{r}

library(dplyr)
table(dat['e0155'])
```

b) Create a new variable that is 0 or 1 depending on whether the children are small or tall (think of a good way how to do this) based on the variable 'groeB' (body height). 
#anyone who's shorter than median value is considerd small
```{r}
#table(dat['groeB'])
summary(dat['groeB'])
dat$abc <- dat$groeB
a <- median(dat$groeB)
a
dat$abc[(dat$groeB <= 136)] <- 0
dat$abc[(dat$groeB >= 136)] <- 1
table(dat$abc)

```
c) Then compute the odds ratio that the mother had hypertension during pregnancy (e0155 == "Ja" (yes), versus e0155 == "Nein" (no)) of tall vs. small children.

```{r, results='asis'}
#options(qwraps2_markup = "markdown")
#to find ratio since we need probability of every comibination i.e. (e0155 == "Ja" (yes), versus e0155 == "Nein" (no)) of tall vs. small children, we are changing the value of differentiation to 100 from 136

dat$xyz <- NA
dat$xyz[(dat$groeB <= 100  & dat$e0155 == "Ja")] <- 0
dat$xyz[(dat$groeB >= 100 & dat$e0155 == "Ja")] <- 1
dat$xyz[(dat$groeB <= 100  & dat$e0155 == "Nein")] <- 2
dat$xyz[(dat$groeB >= 100 & dat$e0155 == "Nein")] <- 3

table(dat$xyz)

height <- c('Small', 'Tall')
hypertension <- c('Ja', 'Nein')
data2 <- matrix(c(198, 2230, 3, 31), nrow=2, ncol=2, byrow=TRUE)
dimnames(data2) <- list('Height'=height, 'Hypertension'=hypertension)

data2
library(epitools)

oddsratio(data2)
```


## Exercise 3 (optional): Confidence intervals

Look at the hypertension variable from exercise 2. Use the binom::binom.confint and the questionr::odds.ratio functions to compute the estimates of the proportion and odds ratio as well as their confidence intervals. (you need to download and load these packages at first).

## Exercise 4 (optional, advanced): Bootstrap

Adapt the bootstrap implementation in R_5b_estimation_bootstrap.Rmd to compute the bootstrap estimate of the standard error of the variance of a normally-distributed and a t-distributed variable. Are they similar? 


