---
title: "Homework 11"
author: "Tauqeer Kasam Rumaney"
date: "January 22, 2023"
output:
  html_document: default
---

Download this R Markdown file, save it on your computer, and perform all the below tasks by inserting your answer in text or by inserting R chunks below. After you are done, upload this file with your solutions on Moodle.

For all exercises, use the KiGGS dataset.

## Exercise 1: Analysis of variance

Choose serum glucose levels (GLUCX) as outcome variable of interest, the factors BMI group (bmiKH) and sex (sex), and physical wellbeing (kw100.e) as quantitative covariate (=metric variable), and compute 5 ANOVA models: (i-ii) one model with each factors separately, (iii) one model with both factors, (iv) one model with both factors and their interaction, (v) one model with both factors, their interaction, and the covariate.

Choose at least one aspect that you observed in the results and write one sentence on the interpretation of this results. For example, you can write about: Write which factor is associated with the outcome? Is there a significant interaction effect? Does anything change after adjusting for the covariate?

```{r}
# Preparation
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06
dat$GLUCX <- as.numeric(as.character(dat$GLUCX))
dat$kw100.e <- as.numeric(as.character(dat$kw100.e))
dat$bmiKH <- factor(dat$bmiKH, labels = c("strong underweight", "underweight", "normal weight", "overweight", "obese"))
dat$sex <- factor(dat$sex, labels = c("boys", "girls"))

```

```{r}
library(car)

leveneTest(dat$GLUCX ~ dat$bmiKH,
  data = dat
)
# since p value is less than 0.05, null hypthesis is rejeted and alternate hypothesis(at least one variance is different) is accepted 

str(dat$bmiKH)

oneway.test(dat$GLUCX ~ dat$bmiKH,
  data = dat,
  var.equal = FALSE 
)
```
```{r}
library(car)

leveneTest(dat$GLUCX ~ dat$sex,
  data = dat
)
# since p value is less than 0.05, null hypthesis is rejeted and alternate hypothesis(at least one variance is different) is accepted 

str(dat$sex)

oneway.test(dat$GLUCX ~ dat$sex,
  data = dat,
  var.equal = TRUE 
)
```
```{r}
res_aov <- aov(dat$GLUCX ~ dat$bmiKH + dat$sex,
  data = dat)
summary(res_aov)
#install.packages("remotes")
#remotes::install_github("easystats/report")
library("report") 

report(res_aov)
```
```{r}
fit1 <- aov(dat$GLUCX ~ dat$bmiKH + dat$sex + dat$bmiKH:dat$sex, data = dat)
summary(fit1)
report(fit1)
```

```{r}
fit2 <- aov(dat$GLUCX ~ dat$bmiKH + dat$sex + dat$kw100.e + dat$bmiKH:dat$sex, data = dat)
summary(fit2)
report(fit2)
```

#Conclusion: The factors interaction has value greater than 0.05, that is null hypothesis is accepted. Also with the covariate addition, interaction doesn't change.


## Exercise 2: Posthoc tests in analysis of variance (optional)

Perform posthoc t-tests with and without a correction for multiple testing for the ANOVA in exercise 1, where we tested whether systolic blood pressure differed between age groups.

```{r}
# Preparation
dat_link <- url("https://www.dropbox.com/s/pd0z829pv2otzqt/KiGGS03_06.RData?dl=1")
load(dat_link)
dat <- KiGGS03_06
dat$age2 <- factor(dat$age2, labels = c("0-1y", "2-3y", "4-5y", "6-7y", "8-9y", "10-11y", "12-13y", "14-15y", "16-17y"))
summary(dat$age2)

#install.packages("multcomp")
library(multcomp)

# Tukey HSD test:
#post_test <- glht(fit1,
  #linfct = mcp(species = "Tukey")
#)
#Error in mcp2matrix(model, linfct = linfct) : Variable(s) ‘species’ have been specified in ‘linfct’ but cannot be found in ‘model’!
#summary(post_test)

```

